Data Flow Analysis\\in the Presence of Correlated Calls

### 2014-12-03 content/datastr.tex
adding comments

It is obvious that two functions represented with the same pairs of sets are equal. Let us prove that two different pairs of sets represent different update functions.

@@

Composing two micro functions means to compose their update-map denotations:
\begin{align}
    \denote{f_1\circ f_2}
    &=\denote{\left\{
              \left(
                r,\,
                  \mathcal N(\langle
                    \ione\cap\itwo,\ 
                    (\ione\cap\utwo)\cup\uone
                  \rangle)
              \right)\,|\,
              r\in\rcc
            \right\}}\notag\\
    &=\footnotemark\lambda m\,.\,\left\{
              \left(
                r,\,
                  \denote{\langle
                    \ione\cap\itwo,\ 
                    (\ione\cap\utwo)\cup\uone
                  \rangle}(m(r))
              \right)\,|\,
              r\in\rcc
            \right\}\notag\\
    &=\lambda m\,.\,\left\{\left(r,\,(\ione\cap m(r)\cap\itwo)\cup(\ione\cap\utwo)\cup\uone\right)\,|\,r\in\rcc\right\}\notag\\
    &=\lambda m\,.\,\left\{\left(r,\,(((m(r)\cap\itwo)\cup\utwo)\cap\ione)\cup\uone\right)\,|\,r\in\rcc\right\}\notag\\
    &=\left(\lambda m\,.\,\{(r,\,(m(r)\cap\ione)\cup\uone)\,|\,r\in\rcc\}\right)\notag\\
      &\quad\circ\left((\lambda m\,.\,\{(r,\,(m(r)\cap\itwo)\cup\utwo)\,|\,r\in\rcc\}\right)\notag\\
    &=\denote{\{(r,\,\setpair{\ione}{\itwo})\}}\circ\denote{\{(r,\,\setpair\itwo\utwo)\}}\notag\\
    &=\denote{f_1}\circ\denote{f_2}.\label{eq:compclosed}
\end{align}
\footnotetext{See Lemma~\ref{lem:normeq}.}
\indent Finally, the meet operation on two micro functions is the union of their update maps:
\begin{align}
    \denote{f_1\sqcap f_2}
    &=\denote{\left\{
              \left(
                r,\,
                  \langle
                    \ione\cup\itwo,\ 
                    \uone\cup\utwo
                  \rangle
              \right)\,|\,
              r\in\rcc
            \right\}}\notag\\
    &=\lambda m\,.\,\left\{\left(r,\,m(r)\cap(\ione\cup\itwo)\cup\uone\cup\utwo\right)\,|\,r\in\rcc\right\}\notag\\
    &=\lambda m\,.\,\left\{\left(r,\,((m(r)\cap\ione)\cup\uone)\cup((m(r)\cap\itwo)\cup\utwo)\right)\,|\,r\in\rcc\right\}\notag\\
    &=\lambda m\,.\,\left\{\left(r,\,(m(r)\cap\ione)\cup\uone\right)\,|\,r\in\rcc\right\}\notag\\
      &\quad\sqcap\lambda m\,.\,\left\{\left(r,\,(m(r)\cap\itwo)\cup\utwo\right)\,|\,r\in\rcc\right\}\notag\\
    &=\denote{f_1}\sqcap\denote{f_2}.\label{eq:meetclosed}
\end{align}

@@

The following Lemma states that the representation of micro functions is efficient according to the definition  of efficiency discussed in Section~\ref{sec:ccdatastr}.
\begin{lemma}\label{lem:efficient}
  The correlated-call representation of a micro function is efficient.
\end{lemma}

>>>>

It is obvious that two functions represented with the same pairs of sets are equal. The following lemma states that two different pairs of sets represent different update functions.

@@

Composing two micro functions means to compose their update-map denotations:\mtodo{Should this be somehow moved to the appendix? Or should we just say ``It can be shown that ...'' and remove the whole computations at all?}
\begin{align}
    \denote{f_1\circ f_2}
    &=\denote{\left\{
              \left(
                r,\,
                  \mathcal N(\langle
                    \ione\cap\itwo,\ 
                    (\ione\cap\utwo)\cup\uone
                  \rangle)
              \right)\,|\,
              r\in\rcc
            \right\}}\notag\\
    &=\footnotemark\lambda m\,.\,\left\{
              \left(
                r,\,
                  \denote{\langle
                    \ione\cap\itwo,\ 
                    (\ione\cap\utwo)\cup\uone
                  \rangle}(m(r))
              \right)\,|\,
              r\in\rcc
            \right\}\notag\\
    &=\lambda m\,.\,\left\{\left(r,\,(\ione\cap m(r)\cap\itwo)\cup(\ione\cap\utwo)\cup\uone\right)\,|\,r\in\rcc\right\}\notag\\
    &=\lambda m\,.\,\left\{\left(r,\,(((m(r)\cap\itwo)\cup\utwo)\cap\ione)\cup\uone\right)\,|\,r\in\rcc\right\}\notag\\
    &=\left(\lambda m\,.\,\{(r,\,(m(r)\cap\ione)\cup\uone)\,|\,r\in\rcc\}\right)\notag\\
      &\quad\circ\left((\lambda m\,.\,\{(r,\,(m(r)\cap\itwo)\cup\utwo)\,|\,r\in\rcc\}\right)\notag\\
    &=\denote{\{(r,\,\setpair{\ione}{\itwo})\}}\circ\denote{\{(r,\,\setpair\itwo\utwo)\}}\notag\\
    &=\denote{f_1}\circ\denote{f_2}.\label{eq:compclosed}
\end{align}
\footnotetext{See Lemma~\ref{lem:normeq}.}
\indent Finally, the meet operation on two micro functions is the union of their update maps:
\begin{align}
    \denote{f_1\sqcap f_2}
    &=\denote{\left\{
              \left(
                r,\,
                  \langle
                    \ione\cup\itwo,\ 
                    \uone\cup\utwo
                  \rangle
              \right)\,|\,
              r\in\rcc
            \right\}}\notag\\
    &=\lambda m\,.\,\left\{\left(r,\,m(r)\cap(\ione\cup\itwo)\cup\uone\cup\utwo\right)\,|\,r\in\rcc\right\}\notag\\
    &=\lambda m\,.\,\left\{\left(r,\,((m(r)\cap\ione)\cup\uone)\cup((m(r)\cap\itwo)\cup\utwo)\right)\,|\,r\in\rcc\right\}\notag\\
    &=\lambda m\,.\,\left\{\left(r,\,(m(r)\cap\ione)\cup\uone\right)\,|\,r\in\rcc\right\}\notag\\
      &\quad\sqcap\lambda m\,.\,\left\{\left(r,\,(m(r)\cap\itwo)\cup\utwo\right)\,|\,r\in\rcc\right\}\notag\\
    &=\denote{f_1}\sqcap\denote{f_2}.\label{eq:meetclosed}
\end{align}

@@

\begin{mdelete}
The following Lemma states that the representation of micro functions is efficient according to the definition  of efficiency discussed in Section~\ref{sec:ccdatastr}.
\end{mdelete}
\begin{lemma}\label{lem:efficient}
  The correlated-call representation of a micro function is efficient.
\end{lemma}

### 2014-12-03 content/eval.tex
started to move proofs to appendix; removing dependency on minted
adding comments

We also implemented a correlated-call transformation from IFDS into IDE problems that consider correlated calls. This transformation is described in Section~\ref{sec:cctrans}. The flow functions can be easily inferred from Section~\ref{sec:edgefnrep}.

@@

\paragraph{IDE-Implementation Correctness}
To test the correctness of the IDE algorithm implementation, we implemented a copy-constant-propagation IDE problem~\cite{sagiv1996precise}. 
In a copy-constant propagation analysis, a variable is considered constant if it is assigned a constant literal or another variable that is also a constant. For example, in a program
\inputMinted{java}{correct.java}
\verb'a' and \verb'b' are considered constant, but \verb'c' and \verb'd' are not (although \verb'd' would be considered constant in linear-constant propagation).

@@

\begin{example}
  Listing~\ref{list:assertions} illustrates the use of the \verb'isSecret' and \verb'notSecret' assertions.
  \begin{listing}[H]
  \inputMinted{java}{assertions.java}
  \caption{Example usage of \texttt{isSecret} and \texttt{notSecret} assertions in regression tests}
  \label{list:assertions}
\end{listing}
\end{example}

@@

A static initializer is invoked only once, before the instance creation of a class or the access of a static member of that class.
Static initializers are invoked lazily by the Java Virtual Machine~\cite{lindholmjava}. This makes finding out at which program point a static initializer is invoked undecidable~\cite{hubert2009soundly}. To account for static initializers in the analysis would require modifying WALA's control-flow supergraph (which does not have edges to static initializers) or using a data-flow analysis for static initialization. Since the primary purpose of the taint-analysis implementation was to test the correlated-call analysis, we did not include a static-initializer analysis in this work.

@@

\begin{listing}
  \inputMinted{java}{interprocRec.java}
  \caption{Inter-procedurally-correlated calls}
  \label{list:interProcRec}
\end{listing}

>>>>

We also implemented a correlated-call transformation from IFDS into IDE problems that consider correlated calls. This transformation is described in Section~\ref{sec:cctrans}.

@@

\begin{mdelete}
\paragraph{IDE-Implementation Correctness}
To test the correctness of the IDE algorithm implementation, we implemented a copy-constant-propagation IDE problem~\cite{sagiv1996precise}.
In a copy-constant propagation analysis, a variable is considered constant if it is assigned a constant literal or another variable that is also a constant. For example, in a program
\inputMinted{java}{correct.java}
\texttt a and \texttt b are considered constant, but \texttt c and \texttt d are not (although \texttt d would be considered constant in linear-constant propagation).

@@

\begin{example}
  Listing~\ref{list:assertions} illustrates the use of the \verb'isSecret' and \verb'notSecret' assertions.
\begin{figure}
  \centering
  \begin{minipage}{\textwidth}
    \inputMinted{java}{assertions.java}
  \end{minipage}
  \caption{Example usage of \texttt{isSecret} and \texttt{notSecret} assertions in regression tests}
  \label{list:assertions}
\end{figure}
\end{example}

@@

A static initializer is invoked only once, before the instance creation of a class or the access of a static member of that class.
Static initializers are invoked lazily by the Java Virtual Machine~\cite{lindholmjava}. This makes finding out at which program point a static initializer is invoked undecidable~\cite{hubert2009soundly}. To account for static initializers in the analysis would require modifying WALA's control-flow supergraph (which does not have edges to static initializers) or using a data-flow analysis for static initialization. Since the primary purpose of the taint-analysis implementation was to test the correlated-call analysis, we did not include a static-initializer analysis in this work.
\end{mdelete}

@@

\begin{figure}
  \centering
  \begin{minipage}{\textwidth}
    \inputMinted{java}{interprocRec.java}
  \end{minipage}
  \caption{Inter-procedurally-correlated calls}
  \label{list:interProcRec}
\end{figure}

### 2014-12-03 content/impl.tex
adding comments

\begin{example}\label{ex:ifdsresult}
  The solution to the taint-analysis IFDS problem $\mathcal P$ in Listing~\ref{list:examplejava} whose exploded supergraph is presented in Figure~\ref{fig:exampleexploded} looks as follows:
  \small\begin{align*}
    \result_\text{IFDS}(\mathcal P)=\{&(\highlight{\textsf{return}_\texttt{secret}}{greyblue},\,\{\texttt a\}),
      &(\highlight{\textsf{start}_\texttt f}{lightsalmonpink},\,\{\texttt s\}),
      &\qquad\qquad(\highlight{\textsf{return}_\texttt f}{lightsalmonpink},\,\{\texttt r,\,\texttt s\}),\\
%      
      &(\highlight{\textsf{call}_\texttt{A.f}}{greyblue},\,\{\texttt a\}),
      &(\highlight{\texttt{if(s==null)}}{lightsalmonpink},\,\{\texttt s\}),
      &\qquad\qquad(\highlight{\texttt{return r}}{lightsalmonpink},\,\{\texttt r,\,\texttt s\}),\\
%
      &(\highlight{\textsf{return}_\texttt{A.f}}{greyblue},\,\{\texttt a,\,\texttt b\}),
      &(\highlight{\textsf{call}_\texttt f}{lightsalmonpink},\,\{\texttt s\}),
      &\qquad\qquad(\highlight{\textsf{end}_\texttt f}{lightsalmonpink},\,\{\texttt r,\,\texttt s\})\}.\\
%
      &(\highlight{\textsf{end}_\texttt{main}}{greyblue},\,\{\texttt a,\,\texttt b\}),&(\highlight{\texttt{return s}}{lightsalmonpink},\,\{\texttt s\}),
  \end{align*}\normalsize
  All other nodes of the control-flow supergraph are mapped to the empty set.
\end{example}

@@

\begin{example}\label{ex:ideresult}
  Converting the IFDS problem $\mathcal P$ from Example~\ref{ex:ifdsresult} into an equivalent IDE problem and solving it will yield the following result:
    \small\begin{align*}
    \result(\transEq(\mathcal P))=\{&((\highlight{\textsf{return}_\texttt{secret}}{greyblue},\,\texttt a),\,\bot),\\
      &((\highlight{\textsf{call}_\texttt{A.f}}{greyblue},\,\texttt a),\,\bot),\\
      &((\highlight{\textsf{return}_\texttt{A.f}}{greyblue},\,\texttt a),\,\bot),\\
      &((\highlight{\textsf{return}_\texttt{A.f}}{greyblue},\,\texttt b),\,\bot),\\
      &\dots\}.\\
    \end{align*}\normalsize
  Suppose that for a pair $(n,\,d)$, where $n\in N^*$ and $d\in D$, there is no corresponding result in ~$\result_\text{IFDS}(\mathcal P)$ (see Example~\ref{ex:ifdsresult}). Then $(n,\,d)$ appears in $\result(\transEq(\mathcal P))$ as $((n,\,d),\,\top)$.
\end{example}

@@

To summarize, Lemma~\ref{lem:sound} shows that the result $\resultCC$ of a correlated-calls analysis is sound since it overapproximates the data flow of all possible concrete execution paths.
We have also shown in Lemma~\ref{lem:subsetifds} that the correlated-calls analysis improves the precision of the original IFDS result $\result_\text{IFDS}$, because the correlated-calls result $\resultCC$ underapproximates an equivalence-IDE result $\resultEq=\result_\text{IFDS}$.
Finally, we showed that a correlated-call transformation to IDE that considers only correlated-call receivers $\rcc$ achieves the same result $\resultCC$ that is obtained when considering all receivers $R$.

>>>>

\begin{mdelete}
\begin{example}\label{ex:ifdsresult}
  The solution to the taint-analysis IFDS problem $\mathcal P$ in Listing~\ref{list:examplejava} whose exploded supergraph is presented in Figure~\ref{fig:exampleexploded} looks as follows:
  \small\begin{align*}
    \result_\text{IFDS}(\mathcal P)=\{&(\highlight{\textsf{return}_\texttt{secret}}{greyblue},\,\{\texttt a\}),
      &(\highlight{\textsf{start}_\texttt f}{lightsalmonpink},\,\{\texttt s\}),
      &\qquad\qquad(\highlight{\textsf{return}_\texttt f}{lightsalmonpink},\,\{\texttt r,\,\texttt s\}),\\
%      
      &(\highlight{\textsf{call}_\texttt{A.f}}{greyblue},\,\{\texttt a\}),
      &(\highlight{\texttt{if(s==null)}}{lightsalmonpink},\,\{\texttt s\}),
      &\qquad\qquad(\highlight{\texttt{return r}}{lightsalmonpink},\,\{\texttt r,\,\texttt s\}),\\
%
      &(\highlight{\textsf{return}_\texttt{A.f}}{greyblue},\,\{\texttt a,\,\texttt b\}),
      &(\highlight{\textsf{call}_\texttt f}{lightsalmonpink},\,\{\texttt s\}),
      &\qquad\qquad(\highlight{\textsf{end}_\texttt f}{lightsalmonpink},\,\{\texttt r,\,\texttt s\})\}.\\
%
      &(\highlight{\textsf{end}_\texttt{main}}{greyblue},\,\{\texttt a,\,\texttt b\}),&(\highlight{\texttt{return s}}{lightsalmonpink},\,\{\texttt s\}),
  \end{align*}\normalsize
  All other nodes of the control-flow supergraph are mapped to the empty set.
\end{example}
\end{mdelete}

@@

\begin{mdelete}
\begin{example}\label{ex:ideresult}
  Converting the IFDS problem $\mathcal P$ from Example~\ref{ex:ifdsresult} into an equivalent IDE problem and solving it will yield the following result:
    \small\begin{align*}
    \result(\transEq(\mathcal P))=\{&((\highlight{\textsf{return}_\texttt{secret}}{greyblue},\,\texttt a),\,\bot),\\
      &((\highlight{\textsf{call}_\texttt{A.f}}{greyblue},\,\texttt a),\,\bot),\\
      &((\highlight{\textsf{return}_\texttt{A.f}}{greyblue},\,\texttt a),\,\bot),\\
      &((\highlight{\textsf{return}_\texttt{A.f}}{greyblue},\,\texttt b),\,\bot),\\
      &\dots\}.\\
    \end{align*}\normalsize
  Suppose that for a pair $(n,\,d)$, where $n\in N^*$ and $d\in D$, there is no corresponding result in ~$\result_\text{IFDS}(\mathcal P)$ (see Example~\ref{ex:ifdsresult}). Then $(n,\,d)$ appears in $\result(\transEq(\mathcal P))$ as $((n,\,d),\,\top)$.
\end{example}
\end{mdelete}

@@

To summarize, Lemma~\ref{lem:sound} shows that the result $\resultCC$ of a correlated-calls analysis is sound since it overapproximates the data flow of all possible concrete execution paths.
We have also shown in Lemma~\ref{lem:subsetifds} that the correlated-calls analysis improves the precision of the original IFDS result $\result_\text{IFDS}$, because the correlated-calls result $\resultCC$ underapproximates an equivalence-IDE result $\resultEq=\result_\text{IFDS}$.
Finally, we showed in Lemma~\ref{lem:onlycorrrec} that a correlated-call transformation to IDE that considers only correlated-call receivers $\rcc$ achieves the same result $\resultCC$ that is obtained when considering all receivers $R$.
